\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-11-09 doc/USER_MANUAL/11_graphics.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

You have also the option to visualize the distributed databases produced
by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
for details). This will create VTK files for each single partition.
You can then use Paraview \urlwithparentheses{www.paraview.org}
to visualized these partitions.

@@

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
large output files. \texttt{MOVIE\_VOLUME} files are saved in the
\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
set, otherwise the velocity field is saved. The look of a movie is
determined by the half-duration of the source. The half-duration should
be large enough so that the movie does not contain frequencies that
are not resolved by the mesh, i.e., it should not contain numerical
noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
of each source in the \texttt{CMTSOLUTION} file is replaced by
\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
\]
\end{quote}

\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
select the appropriate value of 1.1 $\times$ smallest period. As
usual, for a point source one can set \texttt{half duration} in the
\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
in the \texttt{Par\_file} to get the highest frequencies resolved
by the simulation, but for a finite source one would keep all the
\texttt{half durations} as prescribed by the finite source model and
set \texttt{HDUR\_MOVIE} = 0.0.

@@

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{OUTPUT\_FILES} directory. There are several flags in the
main parameter file \texttt{Par\_file} that control the output of
these movie data files, please see section \ref{cha:Main-Parameter} for
more details. Especially, the following parameters affect the output:
\begin{itemize}
\item \texttt{\small SAVE\_DISPLACEMENT}: to save displacement instead of velocity,
\item \texttt{\small NTSTEP\_BETWEEN\_FRAMES}: to set the timesteps between frames,
\item \texttt{\small USE\_HIGHRES\_FOR\_MOVIES}: to save values at all GLL point instead of element edges.
\end{itemize}

@@

When running xspecfem3D with the
\texttt{\small MOVIE\_VOLUME}
flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
specified in the main \texttt{\small Par\_file},
e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.
The output is saved by each processor at the time interval specified
by
\texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
For all domains,
the either the velocity or displacement field is output to files (depending on setting \texttt{SAVE\_DISPLACEMENT}):
{\small
\begin{verbatim}
proc??????_velocity_X_it??????.bin
proc??????_velocity_Y_it??????.bin
proc??????_velocity_Z_it??????.bin
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
proc??????_displ_X_it??????.bin
proc??????_displ_Y_it??????.bin
proc??????_displ_Z_it??????.bin
\end{verbatim}
}
\noindent
For elastic domains, the divergence and curl taken from the velocity (or displacement) field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$ (or $\nabla\cdot{\bf {d}}$ and $\nabla\times{\bf {d}}$), get stored
as well:
{\small
\begin{verbatim}
proc??????_div_it??????.bin
proc??????_curl_X_t??????.bin
proc??????_curl_Y_it??????.bin
proc??????_curl_Z_it??????.bin
\end{verbatim}
}
\noindent
The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and
\texttt{\small proc??????\_curl\_glob\_it??????.bin}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the
\texttt{\small SUPPRESS\_UTM\_PROJECTION}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).

>>>>

You have also the option to visualize the distributed databases produced
by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
in the main parameter file \texttt{DATA/Par\_file} (see Chapter~\ref{cha:Main-Parameter}
for details). This will create VTK files for each single partition.
You can then use Paraview \urlwithparentheses{www.paraview.org}
to visualized these partitions.

@@

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{DATA/Par\_file}. Turning
on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
large output files. \texttt{MOVIE\_VOLUME} files are saved in the
\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
set, otherwise the velocity field is saved. The look of a movie is
determined by the half-duration of the source. The half-duration should
be large enough so that the movie does not contain frequencies that
are not resolved by the mesh, i.e., it should not contain numerical
noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
of each source in the \texttt{CMTSOLUTION} file is replaced by
\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
\]
\end{quote}

\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
select the appropriate value of 1.1 $\times$ smallest period. As
usual, for a point source one can set \texttt{half duration} in the
\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
in the \texttt{DATA/Par\_file} to get the highest frequencies resolved
by the simulation, but for a finite source one would keep all the
\texttt{half durations} as prescribed by the finite source model and
set \texttt{HDUR\_MOVIE} = 0.0.

@@

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{OUTPUT\_FILES} directory. There are several flags in the
main parameter file \texttt{DATA/Par\_file} that control the output of
these movie data files, please see section \ref{cha:Main-Parameter} for
more details. Especially, the following parameters affect the output:
\begin{itemize}
\item \texttt{\small SAVE\_DISPLACEMENT}: to save displacement instead of velocity,
\item \texttt{\small NTSTEP\_BETWEEN\_FRAMES}: to set the timesteps between frames,
\item \texttt{\small USE\_HIGHRES\_FOR\_MOVIES}: to save values at all GLL point instead of element edges.
\end{itemize}

@@

When running xspecfem3D with the
\texttt{\small MOVIE\_VOLUME}
flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
specified in the main \texttt{DATA/Par\_file},
e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.
The output is saved by each processor at the time interval specified
by
\texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
For all domains,
the either the velocity or displacement field is output to files (depending on setting \texttt{SAVE\_DISPLACEMENT}):
{\small
\begin{verbatim}
proc??????_velocity_X_it??????.bin
proc??????_velocity_Y_it??????.bin
proc??????_velocity_Z_it??????.bin
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
proc??????_displ_X_it??????.bin
proc??????_displ_Y_it??????.bin
proc??????_displ_Z_it??????.bin
\end{verbatim}
}
\noindent
For elastic domains, the divergence and curl taken from the velocity (or displacement) field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$ (or $\nabla\cdot{\bf {d}}$ and $\nabla\times{\bf {d}}$), get stored
as well:
{\small
\begin{verbatim}
proc??????_div_it??????.bin
proc??????_curl_X_t??????.bin
proc??????_curl_Y_it??????.bin
proc??????_curl_Z_it??????.bin
\end{verbatim}
}
\noindent
The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and
\texttt{\small proc??????\_curl\_glob\_it??????.bin}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the
\texttt{\small SUPPRESS\_UTM\_PROJECTION}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).

### 2022-11-09 doc/USER_MANUAL/13_changing_the_model.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

To implement your own external tomographic model(s), you must provide
your own external tomography file(s), and choose between two possible
options:\\
 \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
= tomo}, \\
 or for more user control: \\
 \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
= default}, define the negative \texttt{material\_ID} identifier for
each element in the file \texttt{MESH/materials\_file} and use the
following format in the file \texttt{MESH/nummaterial\_velocity\_file}
when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{verbatim}
domain_ID material_ID tomography elastic file_name 1
\end{verbatim}
where: \\
 \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\\
 \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
-1,-2,...), \\
 \indent \texttt{tomography} keyword for tomographic material definition,
\\
 \indent \texttt{elastic} keyword for elastic material definition,
\\
 \indent \texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.\\

The external tomographic model is represented by a grid of points
with assigned material properties and homogeneous resolution along
each spatial direction x, y and z. The ASCII file \texttt{file\_name}
that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
directory, set in the \texttt{Par\_file}. The format of the file,
as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
directory, looks like Figure \ref{fig:tomography_file}, starting with a header information where
\begin{description}
\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
of the initial and final tomographic grid points along the x direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
of the initial and final tomographic grid points along the y direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
of the initial and final tomographic grid points along the z direction
(in the mesh units, e.g., $m$);
\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
spacing between the tomographic grid points along the x, y and z directions,
respectively (in the mesh units, e.g., $m$);
\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
along the spatial directions x, y and z, respectively; \texttt{NX}
is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
y and z directions, respectively;
\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
these values could be the actual limits of the tomographic parameters
in the grid or the minimum and maximum values to which we force the
cut of velocity and density in the model.
\end{description}
After the first four lines, the tomography file \texttt{file\_name} lists the data record where all tomographic
grid points are listed with the corresponding values of \texttt{vp},
\texttt{vs} and \texttt{rho} (and optionally $Q_{p}$ and $Q_{s}$), scanning the grid along the x coordinate
(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
\texttt{END\_Z}, with step of \texttt{SPACING\_Z}).

\noindent
Each data record line provides the velocity model values in a format like:
{\small
\begin{verbatim}
# data record format: purely isotropic model
x y z   vp    vs    rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x y z   vp    vs    rho Qp  Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values,
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
\texttt{constants.h} found in subdirectory \texttt{src/shared/}.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

@@

To use your anisotropic models, you can either set in the \texttt{Par\_file}
the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
models may be superimposed on the mesh based upon the subroutine in
file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
The call to this subroutine is of the form
\begin{verbatim}
call model_aniso(iflag_aniso, rho, vp, vs, &
                 c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26, &
                 c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
generic pre-defined anisotropic models.
\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
$\rho$, $v_{p}$ and $v_{s}$.
\end{description}
Output from the routine consists of the following non-dimensional
model parameters:
\begin{description}
\item [{\texttt{c11,..,c66}}] 21~dimensionalized
anisotropic elastic parameters.
\end{description}
You can replace the \texttt{model\_aniso.f90} file by your own version
\textit{provided you do not change the call structure of the routine},
i.e., the new routine should take exactly the same input and produce
the required relative output.

>>>>

To implement your own external tomographic model(s), you must provide
your own external tomography file(s), and choose between two possible
options:\\
 \indent (1) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
= tomo}, \\
 or for more user control: \\
 \indent (2) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
= default}, define the negative \texttt{material\_ID} identifier for
each element in the file \texttt{MESH/materials\_file} and use the
following format in the file \texttt{MESH/nummaterial\_velocity\_file}
when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{verbatim}
domain_ID material_ID tomography elastic file_name 1
\end{verbatim}
where: \\
 \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\\
 \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
-1,-2,...), \\
 \indent \texttt{tomography} keyword for tomographic material definition,
\\
 \indent \texttt{elastic} keyword for elastic material definition,
\\
 \indent \texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.\\

The external tomographic model is represented by a grid of points
with assigned material properties and homogeneous resolution along
each spatial direction x, y and z. The ASCII file \texttt{file\_name}
that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
directory, set in the \texttt{DATA/Par\_file}. The format of the file,
as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
directory, looks like Figure \ref{fig:tomography_file}, starting with a header information where
\begin{description}
\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
of the initial and final tomographic grid points along the x direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
of the initial and final tomographic grid points along the y direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
of the initial and final tomographic grid points along the z direction
(in the mesh units, e.g., $m$);
\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
spacing between the tomographic grid points along the x, y and z directions,
respectively (in the mesh units, e.g., $m$);
\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
along the spatial directions x, y and z, respectively; \texttt{NX}
is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
y and z directions, respectively;
\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
these values could be the actual limits of the tomographic parameters
in the grid or the minimum and maximum values to which we force the
cut of velocity and density in the model.
\end{description}
After the first four lines, the tomography file \texttt{file\_name} lists the data record where all tomographic
grid points are listed with the corresponding values of \texttt{vp},
\texttt{vs} and \texttt{rho} (and optionally $Q_{p}$ and $Q_{s}$), scanning the grid along the x coordinate
(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
\texttt{END\_Z}, with step of \texttt{SPACING\_Z}).

\noindent
Each data record line provides the velocity model values in a format like:
{\small
\begin{verbatim}
# data record format: purely isotropic model
x y z   vp    vs    rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x y z   vp    vs    rho Qp  Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values,
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.

The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.

@@

To use your own external model, you can set in the \texttt{DATA/Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{setup/constants.h}.

@@

To use your anisotropic models, you can either set in the \texttt{DATA/Par\_file}
the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
models may be superimposed on the mesh based upon the subroutine in
file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
The call to this subroutine is of the form
\begin{verbatim}
call model_aniso(iflag_aniso, rho, vp, vs, &
                 c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26, &
                 c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
generic pre-defined anisotropic models.
\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
$\rho$, $v_{p}$ and $v_{s}$.
\end{description}
Output from the routine consists of the following non-dimensional
model parameters:
\begin{description}
\item [{\texttt{c11,..,c66}}] 21~dimensionalized
anisotropic elastic parameters.
\end{description}
You can replace the \texttt{model\_aniso.f90} file by your own version
\textit{provided you do not change the call structure of the routine},
i.e., the new routine should take exactly the same input and produce
the required relative output.

### 2022-11-09 doc/USER_MANUAL/14_post_processing.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

You do have further options to change this default output behavior,
given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
directory:
\begin{description}
\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
written out in binary format.
\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MAIN}}] Set to \texttt{.true.}
to have only the main process writing out seismograms. This can
be useful on a cluster, where only the main process node has access
to the output directory.
\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
have the seismograms output to \texttt{LOCAL\_PATH} directory specified
in the main parameter file \texttt{DATA/Par\_file}. In this case,
you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
script located in the \texttt{utils/Cluster/lsf/} directory. The usage
of the script would be e.g.:

>>>>

You do have further options to change this default output behavior,
given in the \texttt{DATA/Par\_file}:
\begin{description}
\item [{\texttt{USE\_BINARY\_FOR\_SEISMOGRAMS}}] Set to \texttt{.true.} to have seismograms
written out in binary format.
\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MAIN}}] Set to \texttt{.true.}
to have only the main process writing out seismograms. This can
be useful on a cluster, where only the main process node has access
to the output directory.
\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
have the seismograms output to \texttt{LOCAL\_PATH} directory specified
in the main parameter file \texttt{DATA/Par\_file}. In this case,
you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
script located in the \texttt{utils/Cluster/lsf/} directory. The usage
of the script would be e.g.:
